---
title: Detection of extreme low sea ice cover in the West Antarctica for 10 colonies of Emperor Penguins
author: "Lucas Krüger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#---------- Download NOAA erddap optimum interpolation SST----

#this part of the script is a modification of the scripts in the heatwaveR vignette
#available at https://robwschlegel.github.io/heatwaveR/articles/OISST_preparation.html


```{r, echo =FALSE}
#---------- Download NOAA erdap optimum interpolation SST----

# The packages we will use
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing
library(raster)
library(terra)
library(sf)

```

# The information for the NOAA OISST data

```{r, echo=FALSE}


rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

```

# This function downloads and prepares data based on user provided start and end dates using three different requests, as my PC did not have memory enough to run on a single request.

```{r}


# spatial units to use

shp<-shapefile("D:/Emperors/Colonies_50km_buffer.shp")
plot(shp)

crs=CRS( "+proj=longlat +datum=WGS84 +no_defs") # coordinate reference system CRS
#load antarctic peninsula shapefile to plot it
# this shapefile was cut from high resolution file from the SCAR Antarctic Digital Database, 2023
# Gerrish, L., Ireland, L., Fretwell, P., & Cooper, P. (2023). High resolution vector polygons of the Antarctic coastline (7.7) [Data set]. UK Polar Data Centre, Natural Environment Research Council, UK Research & Innovation. https://doi.org/10.5285/0be5339c-9d35-44c9-a10f-da4b5356840b'

wap<-shapefile("D:/Emperors/Antarctica_WGS84.shp") 


# plot map with shapefile (Fig 1)
shpdf<-data.frame(shp)

ggplot()+
  
  xlab("Longitude")+ylab("Latitude")+
  theme_bw()+geom_polygon(data=wap,aes(x = long, y = lat, group = group))+
    geom_polygon(data=shp,aes(x = long, y = lat, group = group,label=group),fill="lightblue",colour="black")+
  
geom_text(data=shpdf,aes(label=site_id, 
            x=longitude_,
            y=latitude_e),colour="red2",
            size = 3, hjust=0, vjust=-1) +
      coord_quickmap(xlim=c(-130,-65),ylim=c(-76, -59), expand = TRUE, clip = "on")

(shp$site_id)

shp.sf<-st_as_sf(shp)

unique(shp.sf$site_id)

#Extract coordinates and generates a bounding box 

geom<-st_geometry(shp.sf)

bbox_list <- lapply(st_geometry(shp.sf), st_bbox)

#min and max coordinates of each strata as data frame

df01<-data.frame(t(as.data.frame((bbox_list[[1]])))) 
df02<-data.frame(t(as.data.frame((bbox_list[[2]])))) 
df03<-data.frame(t(as.data.frame((bbox_list[[3]])))) 
df04<-data.frame(t(as.data.frame((bbox_list[[4]])))) 
df05<-data.frame(t(as.data.frame((bbox_list[[5]])))) 
df06<-data.frame(t(as.data.frame((bbox_list[[6]])))) 
df07<-data.frame(t(as.data.frame((bbox_list[[7]])))) 
df08<-data.frame(t(as.data.frame((bbox_list[[8]])))) 
df09<-data.frame(t(as.data.frame((bbox_list[[9]]))))
df10<-data.frame(t(as.data.frame((bbox_list[[10]])))) 


```
#download and process data for Bear Peninsula

```{r, echo=FALSE}

unique(shp.sf$site_id[1])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df01$ymin),max(df01$ymax)), 
                                longitude = c(min(df01$xmin),max(df01$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-21")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_BEAR1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_BEAR2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_BEAR3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Bear_Peninsula.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```


#download and process data for Brow nson Islands

```{r, echo=FALSE}

unique(shp.sf$site_id[2])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df02$ymin),max(df02$ymax)), 
                                longitude = c(min(df02$xmin),max(df02$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_BROWN1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_BROWN2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_BROWN3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Brownson_Island.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```

#download and process data for Bryan Coast

```{r, echo=FALSE}

unique(shp.sf$site_id[3])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df03$ymin),max(df03$ymax)), 
                                longitude = c(min(df03$xmin),max(df03$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_BRYAN1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_BRYAN2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_BRYAN3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Bryan_Coast.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```

#download and process data for Cape Gates

```{r, echo=FALSE}

unique(shp.sf$site_id[4])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df04$ymin),max(df04$ymax)), 
                                longitude = c(min(df04$xmin),max(df04$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_CAPE1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_CAPE2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_CAPE3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Cape_Gates.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```

#download and process data for Noville Peninsula

```{r, echo=FALSE}

unique(shp.sf$site_id[5])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df05$ymin),max(df05$ymax)), 
                                longitude = c(min(df05$xmin),max(df05$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_NOV1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_NOV2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_NOV3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Noville_Peninsula.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```

#download and process data for Pfrogner Point

```{r, echo=FALSE}

unique(shp.sf$site_id[6])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df06$ymin),max(df06$ymax)), 
                                longitude = c(min(df06$xmin),max(df06$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_PFRO1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_PFRO2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_PFRO3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Pfrogner_Point.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```

#download and process data for Rothschild Island

```{r, echo=FALSE}

unique(shp.sf$site_id[7])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df07$ymin),max(df07$ymax)), 
                                longitude = c(min(df07$xmin),max(df07$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_ROTH1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_ROTH2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_ROTH3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Rothschild_Island.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```

#download and process data for Smyley Island

```{r, echo=FALSE}

unique(shp.sf$site_id[8])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df08$ymin),max(df08$ymax)), 
                                longitude = c(min(df08$xmin),max(df08$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_SMYL1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_SMYL2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_SMYL3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Smyley_Island.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```

#download and process data for Thurston Glacier

```{r, echo=FALSE}

unique(shp.sf$site_id[9])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df09$ymin),max(df09$ymax)), 
                                longitude = c(min(df09$xmin),max(df09$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_THUR1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_THUR2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_THUR3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Thurston_Glacier.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```


#download and process data for Verdi Inlet

```{r, echo=FALSE}

unique(shp.sf$site_id[10])

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df10$ymin),max(df10$ymax)), 
                                longitude = c(min(df10$xmin),max(df10$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2023-06-20")))


###download and save data


base::system.time(
  OISST_DF1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_DF2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_DF3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_DF1, file = "OISST_VERD1.Rds")
base::saveRDS(OISST_DF2, file = "OISST_VERD2.Rds")
base::saveRDS(OISST_DF3, file = "OISST_VERD3.Rds")

### summarize data per day

DF1<-plyr::ddply(OISST_DF1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF2<-plyr::ddply(OISST_DF2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

DF3<-plyr::ddply(OISST_DF3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

DF12<-rbind(DF1,DF2)
DF123<-rbind(DF12,DF3)

base::saveRDS(DF123, file = "Verdi_Inlet.Rds")

#remove data and release unused memory

rm(list=c("OISST_DF1","OISST_DF2","OISST_DF3","OISST_sub_dl","shp","maxmin","bbox_list",
          "DF1","DF2","DF3","DF12","DF123"))

gc()



```